\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{booktabs}
\usepackage{array}
\usepackage{amsmath}
\usepackage{url}

% Machine-generated figures (see make_permit_types_brief.R). Do NOT hand-edit
% the \input files or this compiled document; rerun the generator instead.
\input{permit_types_numbers.tex}

\title{\vspace{-2.5em}General vs.\ Individual NPDES Permits:\\
Regulatory Framework and Empirical Patterns in ICIS-NPDES}
\author{}
\date{\today}

\begin{document}
\maketitle
\vspace{-2.5em}

\section*{Scope}
This brief distinguishes the Clean Water Act's two NPDES permitting vehicles --
\textbf{individual} and \textbf{general} permits -- on regulatory grounds, then shows how
that distinction is encoded in the ICIS-NPDES (ECHO) bulk data used in this project, and
how it shapes observed enforcement patterns. All data figures are computed by
\path{make_permit_types_brief.R} from the raw ICIS-NPDES files and the existing
enforcement breakdown in \path{output/enforcement_by_permit_type.csv}
(\path{scripts/enforcement_by_permit_type.R}); see \S\ref{sec:repro}.

\section{The regulatory distinction}

\paragraph{Individual permits.} An individual permit is written for one specific
facility. Its effluent limits combine technology-based limits (set nationally, by
industry category) with \textbf{water-quality-based limits derived from the specific
receiving water}, so two dischargers with identical processes can face different
numeric limits depending on where they discharge. Individual permits require a full,
facility-specific application (40 CFR \S\,122.21) and carry facility-specific monitoring
and Discharge Monitoring Report (DMR) obligations. They are the norm for larger, more
complex dischargers -- municipal treatment plants and major industrial sources.

\paragraph{General permits.} A general permit (40 CFR \S\,122.28) is a single template
permit the permitting authority issues once to cover \textbf{many similar, lower-risk
dischargers} at once -- construction and industrial stormwater, concentrated animal
feeding operations (CAFOs), and small or short-lived operations. A facility obtains
coverage by filing a \textbf{Notice of Intent (NOI)} against the master permit, which
substitutes for an individual application; conditions are standardized and monitoring is
lighter or absent. The relationship is one-to-many: one master general permit can cover
hundreds of thousands of individual coverages (\S\ref{sec:data}).

\paragraph{A common misconception: permit term.} Both permit types are capped at the
\emph{same} statutory maximum: 40 CFR \S\,122.46 limits any NPDES permit, individual or
general, to a fixed term not exceeding five years. The practical difference is
administrative, not legal -- reissuing one general permit covers every facility under it
at once, so general permits are almost always issued for the full five years, while an
individual permit's specific term (and whether it is reissued on time or instead runs on
under administrative continuance) depends on that facility's own permitting history
(see \path{docs/data_quirks.md} on the \texttt{ADC} status flag).

\paragraph{Moving between vehicles.} The Director may require any discharger covered by a
general permit to apply for and obtain an individual permit instead (40 CFR
\S\,122.28(b)(3)) -- e.g., if a facility's risk profile turns out not to fit the general
permit's assumptions. Conversely, an owner/operator covered by a general permit may
itself request exclusion and apply for an individual permit within 90 days of the general
permit's publication.

\paragraph{Representative general-permit programs.} EPA's Multi-Sector General Permit
(MSGP) covers stormwater discharges from industrial facilities across roughly thirty
sectors in areas where EPA is the permitting authority; the Construction General Permit
(CGP) covers stormwater from construction sites disturbing $\geq 1$ acre with a pathway to
a water of the US. CAFOs are typically covered under state- or EPA-issued general
permits as well. These are the concrete real-world instances of the ``construction and
industrial stormwater, CAFOs'' category.

\section{How the distinction appears in ICIS-NPDES}
\label{sec:data}

\texttt{PERMIT\_TYPE\_CODE} marks the permitting vehicle on every ICIS-NPDES permit
record. \texttt{NPD} = individual permit; \texttt{GPC} = a facility's coverage under a
general permit; \texttt{NGP} = the general permit itself (the master record, not a
covered facility) -- a small correction to an earlier internal note that had mislabeled
this as \texttt{GPG}, a code that does not appear anywhere in the raw file. The remaining
codes (\texttt{UFT}, \texttt{IIU}, \texttt{APR}, \texttt{SIN}, \texttt{SNN}) are minor,
low-volume variants (state-issued or non-NPDES-equivalent records); their exact
definitions were not independently confirmed against an authoritative EPA reference table
and are not load-bearing for this analysis.

\begin{table}[h]
\centering
\caption{\texttt{PERMIT\_TYPE\_CODE} frequency, two framings: \textbf{Permits} = distinct
\texttt{EXTERNAL\_PERMIT\_NMBR} values (one row per permit); \textbf{Records} = raw rows
in \texttt{ICIS\_PERMITS.csv} (one row per permit \emph{version} -- reissuances and
modifications each add a row).}\label{tab:codes}
\begin{tabular}{l r r}
\toprule
Code & Permits (distinct) & Records (all versions) \\
\midrule
\input{tab_permit_type_codes_rows.tex}
\end{tabular}
\end{table}

The gap between the two columns matters: general-permit \emph{coverages} (\texttt{GPC})
outnumber individual permits by roughly ten to one on either count, but reissuance
behavior differs -- \npdMultiVersionPct\% of individual permits carry more than one
version (a permit keeps its \texttt{NPDES\_ID} across $\sim$5-year reissuances), versus
\gpcMultiVersionPct\% of general-permit coverages, which are more often
\textbf{renumbered} when the master permit is reissued (e.g., an \texttt{NDR100000}-cycle
coverage becomes an \texttt{NDR110000}-cycle coverage under a new ID) rather than
versioned under the same one.

\paragraph{Major status tracks the individual vehicle.} Of \everMajorTotal{} permits ever
flagged ``major'' in any version, \everMajorIndiv{} (\everMajorIndivPct\%) are individual
permits. Restricting analysis to major \emph{and} individual permits, as this project's
facility panel does, is therefore a modest restriction on the major population and a
severe one on the permit universe as a whole (\npdDistinct{} individual permits of
\permitsTotal{} total).

\paragraph{Facility identity and shared IDs.} Individual permits map cleanly to one
physical site with a stable FRS registry identifier (\texttt{FACILITY\_UIN}) and point
geocode. General-permit coverages frequently \textbf{share one generic
\texttt{FACILITY\_UIN}} across unrelated sites: the most extreme case in this data is one
North Dakota UIN (\texttt{\maxGpcUin}) spanning \maxGpcCount{} distinct general-permit
coverages (construction-stormwater projects at different locations, sharing an
administrative facility record rather than a physical one). This makes facility-level
aggregation reliable for individual permits and hazardous for general ones.

\paragraph{Both at once.} \dualPermitFacilities{} \texttt{FACILITY\_UIN}s hold both an
individual (\texttt{NPD}) and a general (\texttt{GPC}) permit. This is typically an
individually permitted process discharge \emph{plus} a general stormwater coverage at the
same site -- the two permits govern different discharges, not the same one, so they
should not be collapsed into a single "one facility, one permit" record.

\paragraph{Data richness.} Monitoring and enforcement are concentrated on the
individual/major side. Per \path{docs/data_quirks.md}, FY2025 DMR data shows majors at
$\sim$84\% reporting coverage versus $\sim$7\% for minors -- consistent with general
permits' lighter-to-absent numeric monitoring requirements.

\section{Enforcement patterns by permit type}

\begin{table}[h]
\centering
\footnotesize
\caption{Formal vs.\ informal enforcement action records, by permit type and
major/minor status (all years, all actions in the bulk enforcement files; each row is one
action record, not one facility).}\label{tab:enforcement}
\begin{tabular}{l l r r r r}
\toprule
Permit type & Status & Formal & Informal & Total & \% Informal \\
\midrule
\input{tab_enforcement_rows.tex}
\end{tabular}
\end{table}

Pooling across major/minor status, individual permits accrue \enfIndivTotal{} action
records (\enfIndivFormalPct\% formal / \enfIndivInformalPct\% informal) and general
permits \enfGenTotal{} (\enfGenFormalPct\% formal / \enfGenInformalPct\% informal). The
two permit types look broadly similar in their formal/informal \emph{split}, but arrive
there by different detection pathways:

\begin{itemize}
  \item \textbf{Individual permits} (mostly majors) generate violations chiefly through
    routine DMR review (numeric effluent-limit exceedances) and scheduled inspections.
    Individual/major permittees are also the population to which EPA's
    \textbf{significant noncompliance (SNC)} criteria are principally applied: since 1995,
    EPA policy has defined SNC to include, among other triggers, any monthly average
    effluent limit exceeded by $\geq$40\% (or by $\geq$20\% for two or more consecutive
    months), any limit violated for four or more months in a two-consecutive-quarter
    window, unauthorized bypass/pass-through, and reporting violations 30+ days late. These
    criteria presuppose numeric monthly DMR limits -- exactly what individual permits
    carry and general permits typically do not -- which is why SNC tracking is
    substantively an individual/major-permit concept rather than a general-permit one.
  \item \textbf{General permits} have less routine numeric monitoring, so violations more
    often surface through inspections, failure to file the NOI or other required paperwork,
    or complaints, rather than a DMR exceedance. Consistent with this, the response is more
    often informal (\enfGenInformalPct\% vs.\ \enfIndivInformalPct\% for individual). The
    Director's authority to revoke general-permit coverage and require an individual permit
    (\S\,1 above; 40 CFR \S\,122.28(b)(3)) is the formal escalation path when a covered facility
    turns out to warrant closer, facility-specific regulation.
\end{itemize}

\section{The 2015 NPDES Electronic Reporting Rule}

EPA's NPDES Electronic Reporting Rule took effect \textbf{December 21, 2015}, replacing
paper-based NPDES reporting with mandatory electronic submission, with the stated goal of
more complete, timely, and nationally consistent compliance data. Implementation was
phased: \textbf{Phase 1} (largely individual-permit DMR data) had a compliance deadline
of December 21, 2016; \textbf{Phase 2} (general-permit and other remaining data streams)
was originally due December 21, 2020 but was later extended by rule, giving affected
programs up to five additional years. This phased rollout is the mechanism behind the
"2016 eRule" reporting break already flagged in this project's data documentation
(\path{docs/data_quirks.md}): majors' DMR coverage was already high before the rule,
while minors' and general-permit coverage only began rising as Phase 1, and later Phase
2, requirements took hold -- so pre- and post-2016 (and pre/post-2020+) coverage are not
directly comparable, particularly for general-permit-covered dischargers.

\section{Implications for analysis}

For facility-level research, the individual permit is the meaningful analytical unit: one
physical site, a stable identifier, dense measurement, and a trustworthy location. The
general permit is best understood as an \textbf{administrative umbrella, not a facility}
-- a single \texttt{FACILITY\_UIN} or master permit number can span hundreds of unrelated
sites, and monitoring/enforcement data is comparatively sparse. Restricting to major
individual permits, as this project's panel does, buys a well-defined and well-measured
population at a real cost to external validity: it excludes the large majority of
permitted dischargers, which are general, lightly monitored, and (per
\S\,3) usually resolved informally rather than through the DMR/inspection/SNC pipeline
that individual permits generate.

\section*{Reproducibility}
\label{sec:repro}
Every figure above is generated by \path{make_permit_types_brief.R} from
\path{data/raw/npdes_downloads/ICIS_PERMITS.csv},
\path{data/raw/npdes_downloads/ICIS_FACILITIES.csv}, and the pre-built
\path{output/enforcement_by_permit_type.csv} (from
\path{scripts/enforcement_by_permit_type.R}, not recomputed here). Regulatory
citations (40 CFR \S\S\,122.28, 122.46; the 1995 SNC criteria memorandum; the 2015
Electronic Reporting Rule and its Phase 2 extension) are drawn from the Federal Register
and eCFR and are not derived from project data. Rerun the generator after any raw-data
refresh; do not hand-edit the \texttt{\textbackslash input} fragments or this PDF.

\end{document}
